---
title: "Conduct Kruskal-Wallis Test for each factor scores chapter two"
author: "Hope & Jeffrey Yarus"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  word_document: 
    toc: yes
    toc_depth: '6'
  html_document: default
  pdf_document: 
    highlight: tango
    latex_engine: xelatex
    number_sections: yes
    toc: yes
    toc_depth: 6
  always_allow_html: yes
---
# Description
This example illustrates hypothesis testing on chemofacies groups derived from XRF elemental data.  It demonstrates how nonparametric statistical tests (e.g., Kruskal–Wallis) are used to evaluate differences in median factor scores and elemental concentrations between identified chemofacies.  The workflow integrates reproducible data processing, visualization, and interpretation steps applied in my research.

# Data loading, loading data that contain the factors and chemofacies
```{r}

# Load necessary libraries
library(readr)
library(dplyr)
library(ggpubr)
#library(dunn.test)
library(rstatix)

# Read the data
data <- read_csv(
  "/mnt/vstor/CSE_MSE_RXF131/staging/sdle/geospatial/core_xrf_xrd/XRF_two_core_in1/PCA_FA_for_two_cores/factorscores_chemofacies.csv"
)

# Set global option for number formatting
options(scipen = 999, digits = 4)

# Explore the data
summary(data)


```



# Calculate median factor scores for each cluster

```{r}
library(dplyr)

# Ensure dplyr functions are used
median_scores <- data %>%
  group_by(chemofacies) %>%
  dplyr::summarise(across(starts_with("Factor"), ~ round(median(.x, na.rm = TRUE), 2)))

print("Median Factor Scores by chemofacies:")
print(median_scores)

```


# Conduct Kruskal-Wallis Test for each factor
```{r}
# Conduct Kruskal-Wallis Test for each factor
kruskal_results <- lapply(data[1:3], function(factor) {
  kruskal.test(factor ~ data$chemofacies)
})

# Display the results with formatted p-values
print("Kruskal-Wallis Test Results:")
for (i in 1:3) {
  factor_name <- names(data)[i]
  p_value <- kruskal_results[[i]]$p.value
  formatted_p <- ifelse(p_value < 0.00001, "< 0.00001", format.pval(p_value, digits = 4))
  cat(
    factor_name,
    ": H-statistic =",
    kruskal_results[[i]]$statistic,
    ", p-value =",
    formatted_p,
    "\n"
  )
}

```


#Visualize the Distribution:


```{r}
# Ensure p-values are formatted as desired
options(scipen = 999, digits = 4)

# Define custom color palette for chemofacies
chemofacies_colors <- c(
  "1" = "#666666",
  "2" = "#CC9933",
  "3" = "#1B9E77",
  "4" = "#7570B3",
  "5" = "#66A61E"
)

# Function to create fancy boxplot
create_plot <- function(element, facies_colors) {
  ggplot(data, aes(
    x = factor(chemofacies),
    y = get(element),
    fill = factor(chemofacies)
  )) +
    geom_boxplot(
      outlier.color = "red",
      outlier.shape = 3,
      color = "black"
    ) +
    stat_boxplot(geom = "errorbar",
                 width = 0.25,
                 color = "black") +
    scale_fill_manual(values = facies_colors) +
    theme_minimal(base_size = 18) +
    labs(title = element, y = element, x = "chemofacies") +
    theme(
      plot.title = element_text(size = 20, face = "bold"),
      axis.title.x = element_text(size = 18),
      axis.title.y = element_text(size = 18),
      axis.text.x = element_text(size = 16),
      axis.text.y = element_text(size = 16),
      legend.position = "none",
      panel.background = element_rect(fill = "grey90", color = NA),
      panel.grid.major.y = element_line(color = "grey80"),
      panel.grid.minor.y = element_blank(),
      panel.grid.major.x = element_blank(),
      plot.background = element_rect(fill = "white", color = NA)
    )
}

# Visualize the distribution of factor scores across clusters
for (i in 1:3) {
  factor_name <- names(data)[i]
  p <- create_plot(factor_name, chemofacies_colors)
  
  # Extract p-value and format it
  kruskal_test <- kruskal.test(as.formula(paste(factor_name, "~ chemofacies")), data = data)
  formatted_p <- ifelse(
    kruskal_test$p.value < 0.00001,
    "Kruskal-Wallis, p < 0.00001",
    paste(
      "Kruskal-Wallis, p =",
      format.pval(kruskal_test$p.value, digits = 4)
    )
  )
  
  # Add the formatted p-value to the plot
  p <- p + annotate(
    "text",
    x = 1.5,
    y = max(data[[factor_name]], na.rm = TRUE) * 0.9,
    label = formatted_p,
    size = 5,
    hjust = 0
  )
  
  print(p)
}

```






