---
title: "Appendix B - Core Analysis;Data Pre-Treatment and Exploration for Chapter Three"
author: ""
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document: 
    highlight: tango
    latex_engine: xelatex
    number_sections: yes
    toc: yes
    toc_depth: 6
  html_document: default
  word_document: 
    toc: yes
    toc_depth: '6'
  always_allow_html: yes
---


# Loading XRD datasets
```{r}
library(readxl)
# Loading the second sheet
XRD_Utica <- read_excel("/mnt/vstor/CSE_MSE_RXF131/staging/sdle/geospatial/core_xrf_xrd/chapter_two/Appendix 1_XRD_Data.xlsx", sheet = 2)
#View(XRD_Utica)
# Loading the third sheet
XRD_PP <- read_excel("/mnt/vstor/CSE_MSE_RXF131/staging/sdle/geospatial/core_xrf_xrd/chapter_two/Appendix 1_XRD_Data.xlsx", sheet = 3)

#View(XRD_PP)

```

#  clean data, get unique names in the specific column from each data frame
```{r}
# Get unique names in the specific column from each data frame
unique_XRD_Utica <- unique(XRD_Utica$API)
unique_XRD_PP <- unique(XRD_PP$API)
```
#Find unique names that are present in both df1 and df2

```{r}
# Find unique names that are present in both df1 and df2
common_names <- intersect(unique_XRD_Utica, unique_XRD_PP)

# Display the results

print(common_names)

```

#This will create two new data frames, XRD_Utica_common and XRD_PP_common, which contain only the rows where the API column matches the common_names. and aslo selecting the county i want
```{r}
library(dplyr)
# Define the county names to remove

## Subset XRD_Utica with common API values
XRD_Utica_common <- XRD_Utica[XRD_Utica$API %in% common_names, ]

# Define the county names to remove
counties_to_remove <- c("HURON", "RICHLAND","LICKING",  "PICKAWAY", "WYANDOT", "MARION", "WARREN")

# Filter out the rows with those county names
XRD_Utica_common_filtered <- XRD_Utica_common %>%
  filter(!COUNTYNM %in% counties_to_remove)

## Subset XRD_PP with common API values
XRD_PP_common <- XRD_PP[XRD_PP$API %in% common_names, ]

# Filter out the rows with those county names
XRD_PP_common_filtered <- XRD_PP_common %>%
  filter(!COUNTYNM %in% counties_to_remove)

# View the new data frames
#View(XRD_PP_common_filtered)
#View(XRD_Utica_common_filtered)
names(XRD_PP_common_filtered)
#names(XRD_PP_common)


```




#To keep only the specified columns in the XRD_Utica_common and XRD_PP_common data frames, you can subset the data frames using the column names you're interested in. 

```{r}
# Define the columns you want to keep
columns_to_keep <- c("API","Quartz", "Calcite","WH_LONG83", "WH_LAT83", "Dolomite", "K-Feldspar", "Plagioclase", "Total Clay",  "Dolomite%ofTotalCarb", "Total Carb.",  "COUNTYNM", "Organic Carbon", "MBI ((W(Q+F+Cal+Dol))/W(tot))")

# Subset XRD_Utica_common with the selected columns
XRD_Utica_common_subset <- XRD_Utica_common_filtered[, columns_to_keep]

# Subset XRD_PP_common with the selected columns
XRD_PP_common_subset <- XRD_PP_common_filtered[, columns_to_keep]

# View the new data frames
#View(XRD_Utica_common_subset)
#View(XRD_PP_common_subset)

```


#To calculate the average of all variables for each unique API in the XRD_Utica_common_subset and XRD_PP_common_subset data frames, you can group the data by the API column and then compute the mean for each group.
```{r}
# Load the necessary library
library(dplyr)

# Calculate the average of all variables except "WH_LONG83", "WH_LAT83", and "COUNTYNM" for each unique API
XRD_Utica_avg <- XRD_Utica_common_subset %>%
  group_by(API) %>%
  dplyr::summarise(
    across(-c(WH_LONG83, WH_LAT83, COUNTYNM), mean, na.rm = TRUE),  # Calculate average for all other variables
    WH_LONG83 = first(WH_LONG83),  # Retain the first value of WH_LONG83
    WH_LAT83 = first(WH_LAT83),    # Retain the first value of WH_LAT83
    COUNTYNM = first(COUNTYNM)     # Retain the first value of COUNTYNM
  )

# View the resulting data frame
#View(XRD_Utica_avg) 


# Calculate the average of all variables except "WH_LONG83", "WH_LAT83", and "COUNTYNM" for each unique API
XRD_PP_avg <- XRD_PP_common_subset %>%
  group_by(API) %>%
  dplyr::summarise(
    across(-c(WH_LONG83, WH_LAT83, COUNTYNM), mean, na.rm = TRUE),  # Calculate average for all other variables
    WH_LONG83 = first(WH_LONG83),  # Retain the first value of WH_LONG83
    WH_LAT83 = first(WH_LAT83),    # Retain the first value of WH_LAT83
    COUNTYNM = first(COUNTYNM)     # Retain the first value of COUNTYNM
  )

# View the resulting data frame
#View(XRD_PP_avg) 


```


#The code is preparing two cleaned datasets (XRD_Utica_no_NA and XRD_PP_no_NA) by removing rows that have missing values in the K-Feldspar column. These cleaned datasets can then be used for further analysis, ensuring that all rows in the K-Feldspar column contain valid data.

```{r}
# Load the necessary library
library(dplyr)

# Subset XRD_Utica_common_subset excluding rows with NA in K-Feldspar
XRD_Utica_no_NA <- XRD_Utica_avg %>%
  filter(!is.na(`K-Feldspar`))

# Subset XRD_PP_common_subset excluding rows with NA in K-Feldspar
XRD_PP_no_NA <- XRD_PP_avg %>%
  filter(!is.na(`K-Feldspar`))

# View the new data frames

#View(XRD_Utica_no_NA)
#View(XRD_PP_no_NA)


```



#Loading TOC DATA
```{r}
library(readxl)
# Loading the second sheet
TOC <- read_excel("/mnt/vstor/CSE_MSE_RXF131/staging/sdle/geospatial/core_xrf_xrd/chapter_two/Appendix 5_TOC.xlsx")

# Filter the rows where Unit column has values "UTCA" or "PNPL"
filtered_TOC <- TOC %>%
  filter(Unit %in% c("UTCA", "PNPL"))

# View the filtered data frame
#View(filtered_TOC)
```



#To create two separate data frames from the common_values data frame based on the values in the Unit column (UTCA and PNPL), 
```{r}
# Load necessary library
library(dplyr)

# Create a data frame for rows where Unit is "UTCA"
common_values_UTCA <- filtered_TOC %>%
  filter(Unit == "UTCA")

# Create a data frame for rows where Unit is "PNPL"
common_values_PNPL <- filtered_TOC %>%
  filter(Unit == "PNPL")

# View the resulting data frames
#View(common_values_UTCA)
#View(common_values_PNPL)

```


# Calculate the average of Organic C% for each unique API Number and keep other variables unchanged
```{r}
# Load necessary library
library(dplyr)

# Calculate the average of Organic C% for each unique API Number and keep other variables unchanged
Utica_avg <- common_values_UTCA %>%
  group_by(`API Number`) %>%
 dplyr::summarise(
    `Organic C%` = mean(`Organic C %`, na.rm = TRUE),
    across(-`Organic C%`, ~ first(.))
  )

# View the resulting data frame
#View(Utica_avg)

```


#Calculate the average of Organic C% for each unique API Number and keep other variables unchanged
```{r}
# Load necessary library
library(dplyr)

# Calculate the average of Organic C% for each unique API Number and keep other variables unchanged
PP_avg <- common_values_PNPL %>%
  group_by(`API Number`) %>%
 dplyr::summarise(
    `Organic C%` = mean(`Organic C %`, na.rm = TRUE),
    across(-`Organic C%`, ~ first(.))
  )

# View the resulting data frame
#View(PP_avg)

```

#Create a new data frame with matching API Number from PP_avg and API from XRD_Utica_no_NA
```{r}
# Load necessary library
library(dplyr)

# Create a new data frame with matching API Number from Utica_avg and API from XRD_Utica_no_NA
Utica_matched_df <- inner_join(Utica_avg, XRD_Utica_avg, by = c("API Number" = "API")) %>%
   dplyr::select(`API Number`, `Organic C%`)

# View the resulting data frame
#View(Utica_matched_df)


```

#Add the Organic C% values from matched_df to XRD_Utica_no_NA based on matching API and API Number
```{r}
# Load necessary library
library(dplyr)

# Add the Organic C% values from matched_df to XRD_Utica_no_NA based on matching API and API Number
XRD_Utica_avg <- XRD_Utica_avg %>%
  left_join(dplyr::select(Utica_matched_df, `API Number`, `Organic C%`), by = c("API" = "API Number"))

# View the updated data frame with the new Organic C% column
#View(XRD_Utica_avg)

```

#Create a new data frame with matching API Number from PP_avg and API from XRD_Utica_no_NA

```{r}
# Load necessary library
library(dplyr)

# Create a new data frame with matching API Number from PP_avg and API from XRD_Utica_no_NA
PP_matched_df <- inner_join(PP_avg, XRD_PP_avg, by = c("API Number" = "API")) %>%
  dplyr::select(`API Number`, `Organic C%`)

# View the resulting data frame
#View(PP_matched_df)

```


#Add the Organic C% values from matched_df to XRD_PP_no_NA based on matching API and API Number
```{r}
# Load necessary library
library(dplyr)

# Add the Organic C% values from matched_df to XRD_Utica_no_NA based on matching API and API Number
XRD_PP_avg <- XRD_PP_avg %>%
  left_join(dplyr::select(PP_matched_df, `API Number`, `Organic C%`), by = c("API" = "API Number"))

# View the updated data frame with the new Organic C% column
#View(XRD_PP_avg)

```

#Get map data for Ohio
```{r}
library(ggplot2)
library(maps)

# Get map data for Ohio
ohio_map <- map_data("county", "ohio")

# Create plot with county names displayed
p <- ggplot() + 
  geom_polygon(data = ohio_map, aes(x = long, y = lat, group = group),
               fill = NA, color = "black") +
  geom_point(data = XRD_PP_avg, aes(x = WH_LONG83, y = WH_LAT83, color = COUNTYNM), size = 2) +
  geom_text(data = XRD_PP_avg, aes(x = WH_LONG83, y = WH_LAT83, label = COUNTYNM),
            vjust = -0.5, hjust = 0.5, size = 3, color = "black") +  # Adjust vjust and hjust to position the labels
  labs(title = "Distribution of Data Points in Ohio", x = "Longitude", y = "Latitude") +
  theme_bw()

# Display plot
p
```



#Ohio_Map_Distribution.png

```{r}
library(ggplot2)
library(dplyr)
library(maps)
library(tools)

# Get map data for Ohio
ohio_map <- map_data("county", "ohio")

# List of counties to include
selected_counties <- c("ASHTABULA", "CARROLL", "COLUMBIANA", "COSHOCTON", "GEAUGA", 
                       "HARRISON", "HOCKING", "NOBLE", "PORTAGE", "STARK", "TRUMBULL", 
                       "TUSCARAWAS", "WASHINGTON")

# Calculate centroids of counties for labels
county_centroids <- ohio_map %>%
  group_by(subregion) %>%
  summarize(long = mean(range(long)), lat = mean(range(lat))) %>%
  filter(toupper(subregion) %in% selected_counties) %>%  # Filter only selected counties
  mutate(subregion = toTitleCase(subregion))  # Capitalize first letter of each word

# Create plot with appropriate projection for Ohio
p <- ggplot() + 
  geom_polygon(data = ohio_map, aes(x = long, y = lat, group = group),
               fill = NA, color = "black") +
  geom_point(data = XRD_PP_avg, aes(x = WH_LONG83, y = WH_LAT83), size = 2, color = "black") + # Plot points
  geom_text(data = county_centroids, aes(x = long, y = lat, label = subregion),
            vjust = 0.5, hjust = 0.5, size = 2.5, color = "black") +  # Use centroids for selected counties
  labs(title = "", x = "Longitude", y = "Latitude") +
  theme_bw() + 
  theme(
    panel.border = element_blank(), # Remove the box or frame around the map
    legend.position = "none"       # Optionally remove the legend
  ) +
  coord_map("mercator")  # Use map projection for proper dimensions

# Save the plot
ggsave(filename = "Ohio_Selected_Counties_Corrected.png", 
       plot = p,       # Use the corrected plot
       height = 6,     # Adjust height
       width = 6,      # Adjust width to keep proportions correct
       dpi = 300,      # High-resolution output
       units = "in")   # Specify units (inches)

# Print the plot
print(p)

```



#To replace all instances of 0.0 with NaN fpr PP
```{r}

XRD_PP_avg <- XRD_PP_avg %>%
  mutate(across(c("K-Feldspar"), ~ na_if(., 0.0)))

# View the updated data frame
#View(XRD_PP_avg)

```

#To replace all instances of 0.0 with NaN for Utica
```{r}
XRD_Utica_avg <- XRD_Utica_avg %>%
  mutate(across(c("K-Feldspar"), ~ na_if(., 0.0)))

# View the updated data frame
#View(XRD_Utica_avg)

```

#for Utica Subset XRD_Utica_avg excluding rows with NA in K-Feldspar and Organic C%
```{r}
# Subset XRD_Utica_avg excluding rows with NA in K-Feldspar and Organic C%
XRD_Utica_no_NA <- XRD_Utica_avg %>%
  filter(!is.na(`K-Feldspar`) & !is.na(`Organic C%`))

# View the updated data frame
#View(XRD_Utica_no_NA)

names(XRD_Utica_no_NA)

# Load necessary libraries
library(dplyr)
library(ggcorrplot)

# Remove "Organic Carbon" and "MBI ((W(Q+F+Cal+Dol))/W(tot))" from the data frame
XRD_Utica_corr_data <- XRD_Utica_no_NA %>%
  dplyr::select(-`Organic Carbon`, -`COUNTYNM`,  -`WH_LONG83`, -`WH_LAT83`, -`API`)

# Calculate the correlation matrix for the remaining numeric columns
corr_matrix <- cor(XRD_Utica_corr_data %>% select_if(is.numeric), use = "complete.obs")



library(corrplot)
corrplot(corr_matrix, method = "circle", tl.cex = 0.8, number.cex = 0.8, addCoef.col = "black")


```


# Want to create two separate linear regression models:
Model 1: Predict Organic C% using Dolomite%ofTotalCarb.
Model 2: Predict K-Feldspar using Dolomite.
Utica
```{r}
# Load necessary library
library(dplyr)

# Model 1: Predict Organic C% using Dolomite%ofTotalCarb
lm_model_organic <- lm(`Organic C%` ~ `Dolomite%ofTotalCarb`, data = XRD_Utica_avg)

# Predict missing values in Organic C%
predicted_organic <- predict(lm_model_organic, newdata = XRD_Utica_avg)

# Replace NA values in Organic C% with the predicted values
XRD_Utica_avg$`Organic C%`[is.na(XRD_Utica_avg$`Organic C%`)] <- predicted_organic[is.na(XRD_Utica_avg$`Organic C%`)]

# Model 2: Predict K-Feldspar using Quartz
lm_model_kfeldspar <- lm(`K-Feldspar` ~ Dolomite, data = XRD_Utica_avg)

# Predict missing values in K-Feldspar
predicted_kfeldspar <- predict(lm_model_kfeldspar, newdata = XRD_Utica_avg)

# Replace NA values in K-Feldspar with the predicted values
XRD_Utica_avg$`K-Feldspar`[is.na(XRD_Utica_avg$`K-Feldspar`)] <- predicted_kfeldspar[is.na(XRD_Utica_avg$`K-Feldspar`)]

# View the updated data frame
#View(XRD_Utica_avg)

# Print summaries for both models
summary(lm_model_organic)
summary(lm_model_kfeldspar)

```

#for PP Subset XRD_Utica_avg excluding rows with NA in K-Feldspar and Organic C%
```{r}
# Subset XRD_Utica_avg excluding rows with NA in K-Feldspar and Organic C%
XRD_PP_no_NA <- XRD_PP_avg %>%
  filter(!is.na(`K-Feldspar`) & !is.na(`Organic C%`))

# View the updated data frame
#View(XRD_PP_no_NA)

names(XRD_PP_no_NA)

# Load necessary libraries
library(dplyr)
library(ggcorrplot)

# Remove "Organic Carbon" and "MBI ((W(Q+F+Cal+Dol))/W(tot))" from the data frame
XRD_PP_corr_data <- XRD_PP_no_NA %>%
  dplyr::select(-`Organic Carbon`, -`COUNTYNM`,  -`WH_LONG83`, -`WH_LAT83`, -`API`)


# Calculate the correlation matrix for the remaining numeric columns
corr_matrix_PP <- cor(XRD_PP_corr_data %>% select_if(is.numeric), use = "complete.obs")



library(corrplot)
corrplot(corr_matrix_PP, method = "circle", tl.cex = 0.8, number.cex = 0.8, addCoef.col = "black")



```

#Predict for PP
```{r}
# Load necessary library
library(dplyr)

# Model 1: Predict Organic C% using Calcite
lm_model_organic_PP <- lm(`Organic C%` ~ Calcite, data = XRD_PP_avg)

# Predict missing values in Organic C%
predicted_organic_PP <- predict(lm_model_organic_PP, newdata = XRD_PP_avg)

# Replace NA values in Organic C% with the predicted values
XRD_PP_avg$`Organic C%`[is.na(XRD_PP_avg$`Organic C%`)] <- predicted_organic_PP[is.na(XRD_PP_avg$`Organic C%`)]

# Model 2: Predict K-Feldspar using Total Carb
lm_model_kfeldspar_PP <- lm(`K-Feldspar` ~ `Plagioclase`, data = XRD_PP_avg)

# Predict missing values in K-Feldspar
predicted_kfeldspar_PP <- predict(lm_model_kfeldspar_PP, newdata = XRD_PP_avg)

# Replace NA values in K-Feldspar with the predicted values
XRD_PP_avg$`K-Feldspar`[is.na(XRD_PP_avg$`K-Feldspar`)] <- predicted_kfeldspar_PP[is.na(XRD_PP_avg$`K-Feldspar`)]

# View the updated data frame
#View(XRD_PP_avg)

# Print summaries for both models
summary(lm_model_organic_PP)
summary(lm_model_kfeldspar_PP)

```

#To add Formation
```{r}
# Load dplyr if not already loaded
library(dplyr)

# Create a new column called Formation and set all values to "Utica"
XRD_PP_avg <- XRD_PP_avg %>%
  mutate(Formation = "PP")

# View the updated data frame
#View(XRD_PP_avg)

# Load dplyr if not already loaded
library(dplyr)

# Create a new column called Formation and set all values to "Utica"
XRD_Utica_avg <- XRD_Utica_avg %>%
  mutate(Formation = "Utica")

# View the updated data frame
#View(XRD_Utica_avg)

```

#Concatenate the two data frames by column names
```{r}
# Load dplyr if not already loaded
library(dplyr)

# Concatenate the two data frames by column names
combined_df <- bind_rows(XRD_Utica_avg, XRD_PP_avg)

# Renaming the column in the data frame
colnames(combined_df)[colnames(combined_df) == "MBI ((W(Q+F+Cal+Dol))/W(tot))"] <- "MBI"

# To check if the column was renamed successfully
head(combined_df)


# View the resulting combined data frame
#View(combined_df)
names(combined_df)
```


#  Carry out PCA  
```{r}
# Load required libraries
library(psych)    # For factor analysis and rotation
library(dplyr)    # For data manipulation
library(pracma)   # For pseudo-inverse function (optional)
library(ggplot2)  # For visualization

# Exclude unwanted columns
excluded_columns <- c("API", "COUNTYNM", "Formation", "Organic Carbon", "WH_LAT83", "WH_LONG83")
FA_df <- combined_df %>%
  dplyr::select(-all_of(excluded_columns))  # Remove non-numeric or irrelevant columns

# Scale the data
FA_scaled <- scale(FA_df)  # Standardize the data (mean=0, sd=1)

# Perform PCA
port.pca <- prcomp(FA_df, center = TRUE, scale = TRUE)  

# Eigenvalues and explained variance
eigenvalues <- port.pca$sdev^2  # Variance explained by each PC
var_explained_percent <- (eigenvalues / sum(eigenvalues)) * 100  # Percentage of variance explained
cum_var_explained_percent <- cumsum(var_explained_percent)  # Cumulative variance explained





```


# Determine Variance Explained Percent

```{r}
library(ggplot2)

# Assuming 'port.pca' is your PCA object
# Calculate the proportion of variance explained by each principal component and convert to percentage
var_explained_percent <- (port.pca$sdev^2 / sum(port.pca$sdev^2)) * 100
cum_var_explained_percent <- cumsum(var_explained_percent)

# Create a data frame for plotting
scree_data <- data.frame(
  PrincipalComponent = seq_along(var_explained_percent),
  VarianceExplainedPercent = var_explained_percent,
  CumulativeVariance = cum_var_explained_percent
)

# Create the Scree plot with individual and cumulative variance
scree_plot <- ggplot(scree_data, aes(x = PrincipalComponent)) +
  geom_bar(aes(y = VarianceExplainedPercent), stat = "identity", fill = "steelblue", alpha = 0.8) +  # Adjust opacity for visibility
  geom_line(aes(y = CumulativeVariance), color = "#1B9E77", size = 1.2, linetype = "solid") +  # Line for cumulative variance
  geom_point(aes(y = CumulativeVariance), color = "#1B9E77", size = 3) +  # Bigger points for cumulative variance
  theme_minimal() +
  labs(
    x = "Principal Component", 
    y = "Percentage of Variance Explained",
    title = "Scree Plot"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),  # Bold and centered title
    axis.title = element_text(size = 14),  # Increase axis labels for readability
    axis.text = element_text(size = 12),  # Improve axis tick size
    panel.grid.major = element_line(color = "grey85", size = 0.3),  # Subtle grid lines
    panel.grid.minor = element_blank()
  ) +
  scale_x_continuous(
    breaks = 1:length(var_explained_percent), 
    labels = paste0("PC", 1:length(var_explained_percent))
  )

# Add labels for individual variance (adjusting vertical spacing)
scree_plot <- scree_plot +
  geom_text(aes(y = VarianceExplainedPercent, label = sprintf("%.1f%%", VarianceExplainedPercent)),
            vjust = -0.5, size = 3, fontface = "bold", check_overlap = TRUE)

# Add labels for cumulative variance (placing slightly above points)
scree_plot <- scree_plot +
  geom_text(aes(y = CumulativeVariance, label = sprintf("%.1f%%", CumulativeVariance)),
            vjust = -0.5, color = "black", size = 3, fontface = "bold", check_overlap = TRUE)

# Save high-resolution image for publication
ggsave("scree_plot.png", plot = scree_plot, dpi = 300, width = 6, height = 4, units = "in")

# Print the Scree plot
print(scree_plot)


```

# Calculate percentage variance explained
```{r}
library(ggplot2)
library(ggfortify)

# Calculate percentage variance explained
var_explained_percent <- (port.pca$sdev^2 / sum(port.pca$sdev^2)) * 100

# Base PCA plot without grouping or color
p <- autoplot(port.pca, loadings = FALSE, colour = I("white")) +
  theme_minimal() +
  labs(
    title = "PCA Biplot",
    x = paste0("Principal Component 1 (", round(var_explained_percent[1], 1), "%)"),
    y = paste0("Principal Component 2 (", round(var_explained_percent[2], 1), "%)")
  ) +
  theme(
    plot.background = element_rect(fill = "white", color = NA),
    panel.background = element_rect(fill = "white", color = NA),
    axis.line = element_line(color = "black", linewidth = 0.5),
    plot.title = element_text(face = "bold", size = 16, hjust = 0.5),
    axis.title = element_text(face = "bold", size = 14),
    axis.text = element_text(size = 12)
  )

# Prepare loadings data (for first 2 PCs)
loadings <- as.data.frame(port.pca$rotation[, 1:2])
names(loadings) <- c("xend", "yend")
loadings$x <- 0
loadings$y <- 0

# Scale arrows for visual clarity
scale_factor <- 0.4
loadings$xend <- loadings$xend * scale_factor
loadings$yend <- loadings$yend * scale_factor

# Add loading vectors to the plot
p_final <- p +
  geom_segment(data = loadings, aes(x = x, y = y, xend = xend, yend = yend),
               arrow = arrow(type = "closed", length = unit(0.1, "inches")),
               color = "#2b2b2b", inherit.aes = FALSE) +
  geom_text(data = loadings, aes(x = xend, y = yend, label = rownames(loadings)),
            vjust = 1.2, hjust = 1.2, size = 3, color = "black", inherit.aes = FALSE)

# Print plot
print(p_final)

# Save high-res output
ggsave("PCA_biplot_PC1_PC2.png", p_final, dpi = 600, width = 8, height = 6, units = "in") 

```



 
 
# Determing the number of factors and  Calculate eigenvalues (Kaiser)

```{r}
# Compute the covariance matrix
covar_matrix <- cov(FA_scaled)

# Calculate eigenvalues
eigenvalues <- eigen(covar_matrix)$values


# Create a bar plot of eigenvalues
barplot(eigenvalues, col = "#1B9E77", xlab = "Index", ylab = "Eigenvalue", border = NA)

# Add a horizontal line at y=1
abline(h = 1.0, col = "red", lty = 2)
```

# Create a barplot for proportion of variance
```{r}
# Compute the covariance matrix
covar_matrix <- cov(FA_scaled)

# Calculate eigenvalues
eigenvalues <- eigen(covar_matrix)$values

# Total variance
total_variance <- sum(eigenvalues)

# Proportion of variance explained
prop_variance <- eigenvalues / total_variance

# Cumulative variance explained
cum_variance <- cumsum(prop_variance)

# Create a barplot for proportion of variance
barplot(prop_variance,
        col = "#1B9E77",
        xlab = "Principal Component",
        ylab = "Proportion of Variance Explained",
        names.arg = paste0("PC", 1:length(prop_variance)),
        ylim = c(0, 1))

# Add a line at the Kaiser criterion (eigenvalue = 1 equivalent)
abline(h = 1 / length(eigenvalues), col = "red", lty = 2)

# Print values
variance_table <- data.frame(
  Component = paste0("PC", 1:length(eigenvalues)),
  Eigenvalue = eigenvalues,
  Proportion = round(prop_variance, 4),
  Cumulative = round(cum_variance, 4)
)

print(variance_table)

```


# Determing the number of factors and  Calculate eigenvalues (Kaiser)

```{r}
# Load required library
library(ggplot2)

# Compute the covariance matrix
covar_matrix <- cov(FA_scaled)

# Calculate eigenvalues and sort them in descending order
eigenvalues <- sort(eigen(covar_matrix)$values, decreasing = TRUE)

# Create a dataframe for ggplot
scree_data <- data.frame(
  Factor = 1:length(eigenvalues),
  Eigenvalue = eigenvalues
)

# Create the refined scree plot
ggplot(scree_data, aes(x = Factor, y = Eigenvalue)) +
  geom_point(size = 3, color = "black") +
  geom_line(color = "black", linewidth = 0.8) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "red", linewidth = 0.8) +
  scale_x_continuous(breaks = scree_data$Factor) +
  scale_y_continuous(breaks = seq(0, max(eigenvalues), by = 1)) +  # NEW: y-axis ticks at 1-unit intervals
  theme_minimal() +
  labs(title = "Scree Plot", x = "Factor Number", y = "Eigenvalue") +
  theme(
    panel.grid.major = element_line(color = "grey90", size = 0.3),
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    axis.title = element_text(size = 14),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5)
  )

# Save the plot
ggsave("scree_plot.png", dpi = 300, width = 6, height = 4, units = "in")





```



# Performing  factor analysis on the entire data sets
```{r}
# Load the psych package for factor analysis functions
library(psych)

# Step 1: Standardize the dataset
# Scaling ensures that all variables contribute equally by converting them to z-scores
FA_scaled <- scale(FA_df)

# Step 2: Test the suitability of the dataset for EFA

# 2a. Kaiser-Meyer-Olkin (KMO) Measure of Sampling Adequacy
# Values > 0.6 suggest that the dataset is suitable for factor analysis
KMO_result <- KMO(FA_scaled)
print(KMO_result)

# 2b. Bartlett’s Test of Sphericity
# Tests whether the correlation matrix significantly differs from the identity matrix
# A p-value < 0.05 indicates that correlations are sufficiently large for EFA
cortest_result <- cortest.bartlett(cor(FA_scaled), n = nrow(FA_scaled))
print(cortest_result)




# Step 3: Determine the number of factors to retain using parallel analysis
# This compares observed eigenvalues with those obtained from random data
# Helps avoid over-extraction based on scree plot alone
fa.parallel(FA_scaled, fa = "fa", n.iter = 100)

# Define number of factors to extract (based on parallel analysis/scree plot)
num_factors <- 2

# Step 4: Run Exploratory Factor Analysis (EFA)
# Performs maximum likelihood extraction of 3 factors from the correlation matrix
# Uses Varimax (orthogonal) rotation to simplify interpretation
efa_result <- fa(r = cor(FA_scaled), 
                 nfactors = 2, 
                 rotate = "varimax", 
                 fm = "ml")

# Step 5: View the rotated factor loading matrix
# Each value represents the strength of association between a variable and a factor
print(efa_result$loadings)

# Step 6: Clean and format the rotated loadings for display

# 6a. Convert loadings to a data frame
rotated_loadings <- as.data.frame(efa_result$loadings[1:nrow(efa_result$loadings), ])

# 6b. Add variable names as a separate column
rotated_loadings_df <- rotated_loadings %>%
  tibble::rownames_to_column("VariableName")

# 6c. Rename columns as Factor1, Factor2, Factor3
colnames(rotated_loadings_df)[-1] <- c("Factor1", "Factor2")

# 6d. Round all factor loadings to 3 decimal places for readability
rotated_loadings_df <- rotated_loadings_df %>%
  mutate(across(starts_with("Factor"), ~ round(.x, 2)))

# 6e. Print the cleaned factor loading table
cat("Rotated Loadings:\n")
print(rotated_loadings_df)

# Step 7: Calculate factor scores for each observation
# These scores represent the position of each sample along the latent factors
efa_scores <- factor.scores(FA_scaled, efa_result)$scores

factor_scores <- as.data.frame(efa_scores)

# Rename factor scores
colnames(factor_scores) <- paste0("Factor", seq_len(num_factors))

# Step 8: Combine factor scores with the original (unscaled) data
# This allows for downstream interpretation and visualization
FA_combined <- cbind(FA_df, factor_scores)

```




#Function to create bar plots for rotated loadings
```{r}
# Function to create bar plots for rotated loadings
create_plot <- function(rotated_loadings_df, indices, filename) {
  # Set up the PNG device with larger dimensions and higher resolution
  png(filename, width = 14, height = 8, units = 'in', res = 400)
  
  # Adjust the layout and margins to fit the plots and prevent the labels from being cut off
  par(mfrow = c(1, length(indices)), mar = c(12, 6, 4, 2), cex.axis = 1.7, cex.lab = 1.7, cex.main = 2.0)
  
  for (i in indices) {
    # Extract the factor name (e.g., "Factor1", "Factor2") based on index
    factor_name <- paste0("Factor", i)
    
    # Extract loadings for the current factor
    loadings_factor <- rotated_loadings_df[[factor_name]]
    
    # Define colors based on the sign of the loadings
    colors <- ifelse(loadings_factor < 0, "red", "blue")
    
    # Plot the loadings with appropriate colors
    barplot(loadings_factor, 
            names.arg = rotated_loadings_df$VariableName,  # Add variable names as labels
            main = factor_name, 
            ylab = "Loadings", 
            col = colors, 
            las = 2,    # Make axis labels perpendicular
            cex.names = 1.3,  # Adjust label size for readability
            space = 0.7)  # Increase space between bars to allow better label fitting
    
    # Add dotted lines to indicate significant loadings
    abline(h = c(-0.5, 0.5), col = "darkgreen", lty = 2)
  }
  
  # Turn off the device
  dev.off()
}

# Create plots for factors using the rotated loadings data frame
create_plot(rotated_loadings_df, indices = 1:2, filename = "factors_PCA_1_2.png")

```


#You can select specific columns from combined_df (e.g., "WH_LAT83", "WH_LONG83", and "Formation") and then add them to your factor_scores_df
```{r}
# Load necessary library
library(dplyr)

# Step 1: Select specific columns from combined_df
selected_columns <- combined_df %>%
  dplyr::select(WH_LAT83, WH_LONG83, Formation)

# Step 2: Add selected columns to factor_scores_df
# Ensure both data frames have the same number of rows
factor_scores_combined <- cbind(factor_scores, selected_columns)

# View the updated data frame
#View(factor_scores_combined)

```
 
 #Create a new data frame for the "Utica" formation &  Create a new data frame for the "PP" formation
 
```{r}
# Load necessary library
library(dplyr)

# Create a new data frame for the "Utica" formation
factor_scores_Utica <- factor_scores_combined %>%
  filter(Formation == "Utica")

# Create a new data frame for the "PP" formation
factor_scores_PP <- factor_scores_combined %>%
  filter(Formation == "PP")

# View the new data frames
#View(factor_scores_Utica)
#View(factor_scores_PP)

# Write the Utica data frame to a CSV file
write.csv(factor_scores_Utica,
          "factor&PCA_scores_Utica.csv",
          row.names = FALSE)

# Write the PP data frame to a CSV file
write.csv(factor_scores_PP, "factor&PCA_scores_PP.csv", row.names = FALSE)


```

